### 2024-12-11 NLP/exercises.tex
Updated some exercises.

\sheet{2023}{1st exam}
\begin{exercise}{Tokenization}
  Give at least 3 examples of ambiguity in tokenization.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Parse trees}
  Given the sentence "John shot the man with the gun" and a grammar: Draw all existing parse trees. Which types of syntactic ambiguity occur here and where?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Stemming and lemmatization}
  Difference between stemming and lemmatization.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Web as corpus}
  List 4 issues you have to deal with when using the web as a corpus for NLP purposes. Explain each.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  How does cross-language QA work? Issues?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  What are the problems with precision, recall, and f-measure in IR? How to improve and why is as an example cohen kappa better?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  Calculate P@5, P@10, MAP for the given IR ranked result list (one for IR system A and one for B. Both go up to ten entries and either have a checkmark or an x in each rank).

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Naive Bayes and SVM}
  Which of the following 4 are correct
  \begin{enumerate}
    \item Features for Naive Bayes classifiers do not need to be uncorrelated
    \item Naive Bayes is slow to train
    \item SVM itself is a multi-label classifier
    \item SVM handles sparse data well
  \end{enumerate}

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus construction}
  How to construct a corpus for linguistic purposes from web data and how to make it usable?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}
  Define hate speech and how to construct a hate speech classifier

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Counter hate speech}
  What is counter hate speech and how to construct such a system

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourced annotation}
  Crowdsourced annotation workflow

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  Why are duplicates on the web and how to detect

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{QA vs IR}
  Differences in QA vs IR

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  Explain applications of multi-language QA and difficulties?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  Explain language detection

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Factual QA vs opinionated QA}
  Differences between factual QA and opinionated QA, and which is easier to evaluate?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{RAG and LangChain}
  What's RAG and how does LangChain help?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Sentiment analysis}
  Discuss: Sentiment analysis using supervised vs. lexicon based approaches.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Edit distance}
  Calculate the Minimum Edit Distance for "TAC" vs. "CAT" (given is also a table just like on the slides, just without the numbers filled in). List all possible word outcomes (in this case 4).

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\sheet{2022}{1st exam}

\begin{exercise}{Corpus construction}
  How would you construct a corpus for linguistic purposes and how to make it usable?

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Sequence tagging and CRF}
  Define sequence tagging and what is CRF

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Minimum edit distance}
  Calculate the Minimum Edit Distance for "TAC" vs. "CAT" (given is also a table just like on the slides, just without the numbers filled in). List all possible word outcomes (in this case 4).

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  Calculate P@5 and MAP for the given IR ranked result list (one for IR system A and one for B. Both go up to ten entries and either have a checkmark or an x in each rank).

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Sequence tagging and CRF}
  Define sequence tagging and what is CRF

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}
  Define hate speech and how to construct a hate speech classifier

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Counter hate speech}
  What is counter hate speech and how to construct such a system

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Sentiment analysis}
  Discuss sentiment detection using lexicon based vs. supervised learning approaches.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Sentiment lexicon extension}
  How could you analyse sentiment with lexicon extension using distributed thesauruses.

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Naive Bayes and SVM}
  Which of the following 4 are correct
  \begin{enumerate}
    \item Features for Naive Bayes classifiers do not need to be uncorrelated
    \item Naive Bayes is slow to train
    \item SVM itself is a multi-label classifier
    \item SVM handles sparse data well
  \end{enumerate}

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  Explain approaches for language detection

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourced annotation}
  Explain the crowdsourced annotation workflow

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{QA vs IR}
  Differences between QA and IR

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  Why are duplicates on the web and how to detect

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  What are the 4 challenges when using the web as a corpus

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

\begin{exercise}{TREC}
  What are the TREC question types

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

>>>>

\sheet{Lecture questions}

\begin{exercise}{Web as an application area for NLP}
  The web is an application area for NLP. Provide examples of applications.

  \begin{solution}
    Examples include:
    \begin{itemize}
        \item Internet of Services
        \item Community mining
        \item Information retrieval
        \item Machine translation for web-based content
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web as a resource for NLP}
  How can the web improve the quality of NLP? Provide examples.

  \begin{solution}
    Examples include:
    \begin{itemize}
        \item Web as Corpus
        \item Analyzing web-based knowledge repositories
        \item Extracting structured information from unstructured web data
        \item Using web-scale data for training language models
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Morphological analysis}
  Identify all stems and affixes (prefix, suffix, infix, circumfix) in the following words: "index," "incorrect," "interesting."

  \begin{solution}
    Morphological breakdown:
    \begin{itemize}
        \item \textbf{index}: stem = "index"
        \item \textbf{incorrect}: prefix = "in," stem = "correct"
        \item \textbf{interesting}: stem = "interest," suffix = "ing"
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Stemming vs lemmatization}
  In contrast to lemmatization, stemming does not necessarily return a valid word form. Why is stemming still useful?

  \begin{solution}
    Stemming is still useful because:
    \begin{itemize}
        \item It is faster and computationally less expensive.
        \item It simplifies implementation for applications like Information Retrieval.
        \item It can still group similar words under a common base, which is often sufficient.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus definition and usage}
  What is a corpus, and how can it be built and used for linguistic analysis?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Corpus definition}: A corpus is a large and structured set of texts, often used for linguistic research.
        \item \textbf{How to build}: Collect text data, preprocess it (cleaning, normalization), and annotate it for specific tasks.
        \item \textbf{Usage}: Analyze linguistic patterns, train NLP models, and evaluate machine learning algorithms.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{NLP datasets}
  What are NLP datasets, and why are they important?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: NLP datasets are collections of annotated or unannotated text data designed for specific NLP tasks (e.g., sentiment analysis, translation).
        \item \textbf{Importance}:
        \begin{itemize}
            \item Serve as a foundation for training and testing machine learning models.
            \item Enable reproducibility in research.
            \item Facilitate comparisons between models and techniques.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus and dataset annotation}
  What is the annotation process for a corpus or dataset, and how does it impact machine learning applications?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Annotation process}: Adding labels or metadata to text data (e.g., part-of-speech tags, sentiment labels, named entities).
        \item \textbf{Impact on machine learning}:
        \begin{itemize}
            \item Improves model performance by providing high-quality training data.
            \item Enables supervised learning by associating input with output labels.
            \item Impacts scalability and generalization of models depending on annotation quality.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Large-scale dataset collection using crowdsourcing}
  What are the main strategies for large-scale dataset collection using crowdsourcing?

  \begin{solution}
    Strategies include:
    \begin{itemize}
        \item \textbf{Task design}: Clearly define annotation tasks with examples and guidelines.
        \item \textbf{Platform selection}: Use platforms like Amazon Mechanical Turk or Prolific.
        \item \textbf{Quality control}: Use redundancy (multiple annotators per task) and include test questions to ensure reliability.
        \item \textbf{Data aggregation}: Combine annotations using majority voting or statistical models.
    \end{itemize}
  \end{solution}
\end{exercise}

% Q: Why are more features not always better for learning a
% classifier?
% A: More training data might be needed; Dependent features
% might lead to deficient models; Some ML algorithms can only deal
% with specific data types; Slower.
% Q: What is the typical type of input data that a sequence tagger
% training step requires?
% A: Labeled training text. (e.g. POS, NER)

\setcounter{section}{2022}
\sheet{1st exam}

\begin{exercise}{Tokenization}
  Give at least 3 examples of ambiguity in tokenization.

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Hyphenated words}: Does "state-of-the-art" count as one token or three?
        \item \textbf{Contractions}: Should "don't" be treated as one token or split into "do" and "n't"?
        \item \textbf{Punctuation}: How should punctuation like periods in "U.S." or apostrophes in possessives ("John's book") be handled?
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Parse trees}
  "\textit{John shot the man with the gun}". Draw all possible parse trees for this sentence using the English grammar. What types of syntactic ambiguity are present?

  \begin{solution}
    A parse tree can be drawn using the following structures: $S, N, V, Det, P, NP, VP, PP$. This is an example of attachment ambiguity. The prepositional phrase "with the gun" can attach to either "shot" or "the man".
  \end{solution}
\end{exercise}

\begin{exercise}{Stemming and lemmatization}
  What is the difference between stemming and lemmatization?

  \begin{solution}
    While stemming reduces a word to its base form by removing pre- and suffices ("running" → "run"), lemmatization considers the word's context and grammar to find its dictionary form ("better" → "good"). Lemmatization is more precise than stemming but computationally more expensive.
  \end{solution}
\end{exercise}

\begin{exercise}{Web as corpus}
  List 4 issues you have to deal with when using the web as a corpus for NLP purposes and explain each.

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Noise}: Content may contain errors, irrelevant information, or non-standard grammar.
        \item \textbf{Duplicates}: Identical or nearly identical data appears multiple times.
        \item \textbf{Bias}: Web content often reflects societal biases, which can skew analyses.
        \item \textbf{Reproducability}: Web pages change frequently, making reproducibility challenging.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  How does cross-language QA work and what are the main issues?

  \begin{solution}
    Cross-language QA involves answering questions in one language based on information in another. This is accomplished by translating the question, retrieving relevant information in the target language, and generating an answer. Challenges include translation errors, cultural differences.
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Data Preprocessing $\to$ Feature extraction $\to$ Model training $\to$ Evaluation $\to$ Deployment
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  What are the problems with precision, recall, and F-measure in IR? How can these be improved? Why is Cohen's Kappa better?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Precision}: How relevant are the retrieved documents? Favors retrieving fewer documents.
        \item \textbf{Recall}: How many of the relevant documents were retrieved? Favors retrieving more documents.
        \item \textbf{F-measure}: Harmonic mean of precision and recall. Balances the two metrics but not adjustable.
        \item \textbf{Cohen's Kappa}: Measures inter-annotator agreement, accounting for chance agreement. Better for evaluating IR systems because it considers agreement beyond chance.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  Calculate P@5, P@10, and MAP for the given IR ranked result list. Analyze both IR system A and system B.

  \begin{solution}
    For P@5, P@10, and MAP:
    \begin{itemize}
        \item \textbf{P@5}: Count relevant documents in the top 5 results and divide by 5.
        \item \textbf{P@10}: Same calculation for the top 10 results.
        \item \textbf{MAP}: Average Precision for each query, then average across all queries.
    \end{itemize}
    Steps:
    \begin{itemize}
        \item Identify positions of relevant documents in ranked lists.
        \item Calculate precision at each relevant position.
        \item Take mean of these precision values.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Naive Bayes and SVM}
  Which of the following statements are correct?
  \begin{enumerate}
    \item Features for Naive Bayes classifiers do not need to be uncorrelated.
    \item Naive Bayes is slow to train.
    \item SVM itself is a multi-label classifier.
    \item SVM handles sparse data well.
  \end{enumerate}

  \begin{solution}
    Correct statements:
    \begin{enumerate}
        \item \textbf{True}: Features for Naive Bayes do not need to be uncorrelated, though it assumes independence for computation.
        \item \textbf{False}: Naive Bayes is fast to train, not slow.
        \item \textbf{False}: SVM is not inherently multi-label but can be adapted for multi-label classification.
        \item \textbf{True}: SVM handles sparse data well due to its kernel functions.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Corpus construction}
  How can a corpus be constructed for linguistic purposes from web data, and how can it be made usable?

  \begin{solution}
    Steps for constructing a corpus:
    \begin{enumerate}
        \item \textbf{Data collection}: Crawl web pages, scrape data, or gather domain-specific text.
        \item \textbf{Cleaning}: Remove duplicates, irrelevant content, and normalize text.
        \item \textbf{Annotation}: Add metadata or labels for tasks like classification or parsing.
        \item \textbf{Usability}: Store in accessible formats (e.g., JSON, XML) and document the corpus for ease of use.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Challenges in pragmatics and discourse analysis:
    \begin{itemize}
        \item \textbf{Context dependency}: Understanding depends heavily on prior sentences or external knowledge.
        \item \textbf{Ambiguity}: Words and phrases can have multiple meanings depending on the situation.
        \item \textbf{Social norms}: Interpretations vary based on cultural or societal factors.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}
  Define hate speech and explain how to construct a hate speech classifier.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Hate speech involves language that attacks or discriminates against individuals or groups based on attributes like race, religion, ethnicity, gender, or sexual orientation.
        \item \textbf{Constructing a hate speech classifier}:
        \begin{enumerate}
            \item Collect labeled datasets with examples of hate speech and non-hate speech.
            \item Preprocess text (e.g., tokenization, removing stop words).
            \item Extract features (e.g., bag-of-words, TF-IDF, embeddings).
            \item Train a supervised machine learning model (e.g., Naive Bayes, SVM, or deep learning models like LSTMs or transformers).
            \item Evaluate the classifier using metrics like precision, recall, and F1-score.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Counter hate speech}
  What is counter hate speech, and how can a system be constructed for it?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Counter hate speech is content designed to reduce the impact of hate speech, often by promoting positive, empathetic, or factual responses.
        \item \textbf{Constructing a counter hate speech system}:
        \begin{enumerate}
            \item Collect datasets of effective counter hate responses.
            \item Train models to generate or retrieve appropriate responses based on the hate speech content.
            \item Use sentiment analysis and empathy detection to craft responses.
            \item Deploy the system in platforms like social media or forums, ensuring ethical considerations are addressed.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourced annotation}
  What is the workflow for crowdsourced annotation?

  \begin{solution}
    Steps in a crowdsourced annotation workflow:
    \begin{enumerate}
        \item \textbf{Task design}: Define the annotation task clearly with examples and instructions.
        \item \textbf{Platform setup}: Use a crowdsourcing platform (e.g., Amazon Mechanical Turk, Prolific).
        \item \textbf{Quality control}: Include test questions and redundancy (multiple annotators per item) to ensure data reliability.
        \item \textbf{Data aggregation}: Aggregate annotations using majority voting or weighted schemes.
        \item \textbf{Evaluation}: Analyze annotation quality and address inconsistencies.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  Why do duplicates exist on the web, and how can they be detected?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Duplicates}: Duplicates arise from mirrored websites, copied content, or redundant crawls.
        \item \textbf{Detection methods}:
        \begin{itemize}
            \item Hashing (e.g., MD5 or SHA): Compare hash values of documents.
            \item Similarity measures: Use cosine similarity or Jaccard index on n-grams.
            \item Structural analysis: Detect similarity in HTML structure.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{QA vs IR}
  What are the differences between Question Answering (QA) and Information Retrieval (IR)?

  \begin{solution}
    Differences between QA (Question Answering) and IR (Information Retrieval):
    \begin{itemize}
        \item \textbf{Goal}:
        \begin{itemize}
            \item QA aims to provide direct answers.
            \item IR retrieves documents containing relevant information.
        \end{itemize}
        \item \textbf{Processing}:
        \begin{itemize}
            \item QA requires deeper natural language understanding.
            \item IR focuses on keyword matching and ranking.
        \end{itemize}
        \item \textbf{Applications}:
        \begin{itemize}
            \item QA: Chatbots, customer support.
            \item IR: Search engines.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Cross-language QA}
  What are the applications of multi-language QA, and what difficulties arise?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Applications}: Multi-language QA systems are used in global customer support, cross-border e-commerce, and multi-language search engines.
        \item \textbf{Difficulties}:
        \begin{itemize}
            \item Translation errors can distort queries or results.
            \item Cultural and linguistic differences complicate understanding and generating answers.
            \item Lack of high-quality bilingual corpora for training.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  How does language detection work?

  \begin{solution}
    Approaches for language detection:
    \begin{enumerate}
        \item \textbf{N-gram analysis}: Compare n-grams of text against language-specific frequency profiles.
        \item \textbf{Machine learning models}: Use classifiers trained on labeled text for different languages.
        \item \textbf{Lexicon-based methods}: Match words in the text to dictionaries of different languages.
        \item \textbf{Hybrid approaches}: Combine multiple methods for higher accuracy.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Factual QA vs opinionated QA}
  What are the differences between factual QA and opinionated QA, and which is easier to evaluate?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Factual QA}: Deals with objective information (e.g., "What is the capital of France?").
        \item \textbf{Opinionated QA}: Involves subjective answers (e.g., "Is Paris a good travel destination?").
        \item \textbf{Evaluation}: Factual QA is easier to evaluate as answers can be verified against ground truth. Opinionated QA requires subjective or aggregated assessments.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{RAG and LangChain}
  What is Retrieval-Augmented Generation (RAG), and how does LangChain assist in implementing it?

  \begin{solution}
    \begin{itemize}
        \item \textbf{RAG (Retrieval-Augmented Generation)}: Combines retrieval models with generative models to produce context-aware answers. External knowledge is retrieved and fed into the generation model.
        \item \textbf{LangChain}: A framework for building applications with large language models, offering tools for integrating RAG systems. It manages chaining retrieval and generation steps efficiently.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Sentiment analysis}
  Compare sentiment analysis using supervised approaches and lexicon-based approaches.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Supervised learning}: Uses labeled data to train machine learning models.
        \begin{itemize}
            \item \textbf{Advantages}: High accuracy with large datasets.
            \item \textbf{Drawbacks}: Relies on annotated data and is domain-dependent.
        \end{itemize}
        \item \textbf{Lexicon-based approaches}: Use sentiment dictionaries to determine polarity.
        \begin{itemize}
            \item \textbf{Advantages}: Simplicity and no training requirement.
            \item \textbf{Drawbacks}: Limited contextual understanding and poor performance on complex text.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Edit distance}
  Calculate the Minimum Edit Distance for "TAC" vs. "CAT" and list all possible word outcomes.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Minimum Edit Distance for "TAC" and "CAT"}:
        \[
        \begin{array}{c|c|c|c|c}
            &  & C & A & T \\
            \hline
            & 0 & 1 & 2 & 3 \\
            T & 1 & 1 & 2 & 2 \\
            A & 2 & 2 & 1 & 2 \\
            C & 3 & 2 & 2 & 2 \\
        \end{array}
        \]
        \item \textbf{Word outcomes}: Possible outcomes from edits:
        \begin{enumerate}
            \item CAT
            \item TAT
            \item TAC
            \item CAC
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\setcounter{section}{2021}
\sheet{1st exam}

\begin{exercise}{Corpus construction}
  How would you construct a corpus for linguistic purposes and how can it be made usable?

  \begin{solution}
    \begin{enumerate}
        \item \textbf{Data collection}: Gather data from relevant sources (e.g., web crawling, domain-specific text).
        \item \textbf{Cleaning and preprocessing}: Remove duplicates, non-linguistic content, and normalize text (e.g., case normalization, tokenization).
        \item \textbf{Annotation}: Add metadata or linguistic labels (e.g., POS tags, syntactic trees) to the text.
        \item \textbf{Storage}: Store the corpus in an accessible format like JSON, XML, or relational databases.
        \item \textbf{Documentation}: Provide guidelines and metadata about the corpus for user understanding.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Context dependency}: Requires understanding of the broader context beyond sentence boundaries.
        \item \textbf{Ambiguity}: Words or phrases can have multiple meanings depending on the discourse.
        \item \textbf{Cultural and social norms}: Interpretation is influenced by societal and cultural factors.
        \item \textbf{Incomplete information}: Missing parts of conversations or text make it harder to analyze.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Sequence tagging and CRF}
  Define sequence tagging and explain what a Conditional Random Field (CRF) is.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Sequence tagging}: The process of assigning labels to each element in a sequence (e.g., part-of-speech tagging, named entity recognition).
        \item \textbf{Conditional Random Field (CRF)}: A probabilistic model used for structured prediction in sequence data. It considers the context of neighboring labels to improve accuracy.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Minimum edit distance}
  Calculate the Minimum Edit Distance for "TAC" vs. "CAT" (a table is given). List all possible outcomes.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Minimum Edit Distance Table}:
        \[
        \begin{array}{c|c|c|c|c}
            &  & C & A & T \\
            \hline
            & 0 & 1 & 2 & 3 \\
            T & 1 & 1 & 2 & 2 \\
            A & 2 & 2 & 1 & 2 \\
            C & 3 & 2 & 2 & 2 \\
        \end{array}
        \]
        \item \textbf{Outcomes}:
        \begin{enumerate}
            \item CAT
            \item TAT
            \item TAC
            \item CAC
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{IR evaluation}
  Calculate P@5 and MAP for given ranked IR results for two systems.

  \begin{solution}
    \begin{itemize}
        \item \textbf{P@5}: Count relevant documents in the top 5 results and divide by 5.
        \item \textbf{MAP}: For each query, calculate average precision across relevant documents and average these scores.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Hate speech}
  Define hate speech and how to construct a hate speech classifier.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Hate speech targets individuals or groups based on attributes like race, religion, gender, or orientation.
        \item \textbf{Classifier construction}:
        \begin{enumerate}
            \item Collect labeled datasets of hate and non-hate speech.
            \item Preprocess data by removing noise and tokenizing.
            \item Train a machine learning model (e.g., Naive Bayes, SVM, transformers).
            \item Evaluate using metrics like F1-score, precision, and recall.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Counter hate speech}
  What is counter hate speech, and how can a system be constructed for it?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Definition}: Counter hate speech neutralizes hate content by promoting positive or factual responses.
        \item \textbf{System construction}:
        \begin{enumerate}
            \item Collect examples of effective counter hate responses.
            \item Use NLP techniques to classify hate speech and generate counter-responses.
            \item Train models on empathy and sentiment analysis for response generation.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Sentiment lexicon extension}
  How can sentiment analysis be improved using lexicon extensions from distributed thesauruses?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Lexicon extension}: Expand sentiment dictionaries using related terms from distributed word embeddings like Word2Vec or GloVe.
        \item \textbf{Process}:
        \begin{enumerate}
            \item Identify synonyms, antonyms, or similar words in embedding spaces.
            \item Update lexicons with these terms to improve coverage.
            \item Validate new terms to ensure sentiment accuracy.
        \end{enumerate}
    \end{itemize}
  \end{solution}
\end{exercise}
\begin{exercise}{Sentiment analysis}
  Discuss sentiment detection using lexicon based vs. supervised learning approaches.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Lexicon-based approaches}: 
        \begin{itemize}
            \item Use pre-built dictionaries of words with sentiment labels (e.g., positive, negative).
            \item Advantages:
                \begin{itemize}
                    \item Easy to implement and interpret.
                    \item No need for labeled training data.
                \end{itemize}
            \item Disadvantages:
                \begin{itemize}
                    \item Limited to the coverage of the lexicon.
                    \item Poor performance on context-dependent sentiment (e.g., sarcasm).
                \end{itemize}
        \end{itemize}
        \item \textbf{Supervised learning approaches}:
        \begin{itemize}
            \item Train machine learning models on labeled data to classify sentiment.
            \item Advantages:
                \begin{itemize}
                    \item Can learn from context and handle subtle sentiment expressions.
                    \item Scalable to large datasets.
                \end{itemize}
            \item Disadvantages:
                \begin{itemize}
                    \item Requires annotated data, which can be costly to obtain.
                    \item More computationally intensive than lexicon-based approaches.
                \end{itemize}
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Context-dependence}: Lexicons fail to account for how sentiment changes with context.
        \item \textbf{Negation handling}: Sentences like "not bad" may be incorrectly classified without understanding negation.
        \item \textbf{Ambiguity}: Words can have different sentiments in different contexts (e.g., "hot" in weather vs. arguments).
        \item \textbf{Domain-specific issues}: Lexicons often do not cover domain-specific terms or slang.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Naive Bayes and SVM}
  Which of the following 4 are correct:
  \begin{enumerate}
    \item Features for Naive Bayes classifiers do not need to be uncorrelated.
    \item Naive Bayes is slow to train.
    \item SVM itself is a multi-label classifier.
    \item SVM handles sparse data well.
  \end{enumerate}

  \begin{solution}
    \begin{itemize}
        \item \textbf{Correct}: 
        \begin{itemize}
            \item Features for Naive Bayes classifiers do not need to be uncorrelated.
            \item SVM handles sparse data well.
        \end{itemize}
        \item \textbf{Incorrect}:
        \begin{itemize}
            \item Naive Bayes is actually fast to train.
            \item SVM is not inherently multi-label but can be adapted for multi-label classification.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  Explain approaches for language detection.

  \begin{solution}
    \begin{itemize}
        \item \textbf{N-gram analysis}: Compare the frequency of character n-grams against language-specific profiles.
        \item \textbf{Machine learning models}: Train classifiers (e.g., SVM, neural networks) using labeled text data for various languages.
        \item \textbf{Lexicon-based methods}: Use dictionaries to match input text to language-specific words.
        \item \textbf{Hybrid approaches}: Combine n-gram, lexicon, and machine learning techniques for better accuracy.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourced annotation}
  Explain the crowdsourced annotation workflow.

  \begin{solution}
    Steps in a crowdsourced annotation workflow:
    \begin{enumerate}
        \item \textbf{Task design}: Define the annotation task with clear instructions and examples.
        \item \textbf{Platform setup}: Deploy the task on crowdsourcing platforms like Amazon Mechanical Turk.
        \item \textbf{Quality control}: Use redundancy by assigning multiple annotators to the same task and include test questions.
        \item \textbf{Data aggregation}: Combine annotations using majority voting or weighted scoring methods.
        \item \textbf{Evaluation}: Analyze the consistency and reliability of annotations.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Ambiguities in tokenization include:
    \begin{itemize}
        \item \textbf{Hyphenation}: Is "state-of-the-art" one token or three?
        \item \textbf{Punctuation}: Should periods in abbreviations (e.g., "U.S.") be separated?
        \item \textbf{Contractions}: Should "don't" be split into "do" and "n't"?
        \item \textbf{Language differences}: Tokenization rules vary by language (e.g., Chinese characters vs. English words).
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{QA vs IR}
  What are the differences between QA and IR?

  \begin{solution}
    Differences between QA and IR:
    \begin{itemize}
        \item \textbf{Goal}:
        \begin{itemize}
            \item QA provides direct answers to user queries.
            \item IR retrieves documents that may contain relevant information.
        \end{itemize}
        \item \textbf{Processing}:
        \begin{itemize}
            \item QA requires natural language understanding and reasoning.
            \item IR focuses on keyword matching and ranking algorithms.
        \end{itemize}
        \item \textbf{Applications}:
        \begin{itemize}
            \item QA is used in chatbots and virtual assistants.
            \item IR powers search engines like Google.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Challenges in cross-language QA include:
    \begin{itemize}
        \item \textbf{Translation errors}: Inaccurate translations can distort queries or answers.
        \item \textbf{Lack of parallel resources}: Few high-quality bilingual corpora exist for certain language pairs.
        \item \textbf{Cultural differences}: Questions or answers may need cultural adaptation.
        \item \textbf{Mismatched structures}: Linguistic differences make aligning queries and answers difficult.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Stemming}: Reduces a word to its root by chopping off affixes (e.g., "running" → "run"). It is less precise and may produce non-words.
        \item \textbf{Lemmatization}: Converts a word to its base form (lemma) based on its context (e.g., "better" → "good"). It is more accurate but computationally expensive.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  Why are duplicates on the web, and how can they be detected?
